---
title: "Patient overview"
author: "Johann Hawe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(knitr)
source("scripts/lib.R")
```

## Preprocessing

In this markdown we merely process the patient data and generate some simple
exploratory plots.
From each of the excel files we have at hand, we extract the information from 
three sheets overall: 'Stammdaten', 'Euroscore II' and 'Verlegungskriterien'.
For each entry we create two distinct keys:

* **key_visit**: A unique identifier for each entry: Name-Surname-Birthday-Entryday
* **key_patient**: A unique identifier for each patient: Name-Surname-Birthday

The **key_visit** identifier is unique over all entries in the result table, whereas
the **key_patient** identifier is only unique for patients, not however for the different 
visits (i.e. a patient being in for surgery more than once has the same **key_patient** but a different **key_visit** for each of the stays).

In the next three sections we briefly describe the general strategies used to extract data and merged them into a single table.

### Stammdaten
We extract most of the data from this sheet 'as is', i.e. we directly use the values
in the individual fields for the entries in our data matrix.
For the more complex data, i.e. "Diagnose_Behandlung", "Anamnese", "Medikation" etc. we always 
extract all the entries below the respective captions. Each entry was then simply concatenated to form a single string (separator used: ';'),
and all entries were in the end merged to a single value for the respective entry / patient (separator used: '|').
All occurrences of the used separators were replaced with "," before concatenating the entries.

### Euroscore II
Again, we just extract the score 'as is', in this case meaning that we create the names
(e.g. Komorbiditaet.IDDM) and add as a value the entry in the respective field of the 'C' column (e.g. 'y').
Finally, we directly extract the risk of martality score from field D47.

### Verlegungskriterien
We create as for the Euroscore II sheet the names for each of the entires, e.g. CVS.Katecholamintherapie.
As values for each of the names we always create triple of the form "JA:;TEILWEISE:;NEIN:", and we
write to each value of the triple the value in the respective field in the sheet, e..g "JA:;TEILWEISE:;NEIN:x".
This was done to avoid creating a multitude of columns (i.e. CVS.Katecholamintherapie.JA, CVS.Katecholamintherapie.TEILWEISE, CVS.Katecholamintherapie.NEIN etc.) but nevertheless contains all the information present in the sheet. In addition,
we extracted the last three rows (43-45) using the same concept but with different triples (i.e. "empfohlen:;erfolgt:;nicht erfolgt:").
Likely, this does not make sense for the last two rows as indicated in the sheet, but is currently used for convenience and can be ignored during any further analysis.
The above mentioned approach can easily be adjusted (e.g. to multiple columns) upon further discussion.

```{r load_patients, echo=F}
patient_dirs <- list.files("results/patient_data/", 
                           pattern="patient_*", full.names = T)
print(paste0("Processing ", length(patient_dirs), " directories with patient data."))

# iterate over all patient directories (containing the extracted csv files per patient)
patient_list <- lapply(patient_dirs, function(pd) {
  print(paste0("Processing directory ", pd, " ."))
  base_data <- process_basedata(paste0(pd, "/Stammdaten.csv"))  
  euroscore_data <- process_euroscore(paste0(pd, "/EUROSCORE II.csv"))
  mvt_criteria <- process_mvt_criteria(paste0(pd, "/Verlegungskriterien.csv"))  
  result <- c(base_data,euroscore_data, mvt_criteria)
  result
})

# create final table
tab <- do.call(rbind, patient_list)

print("First 5 columns and rows of the data:")
print(tab[1:min(5, nrow(tab)),1:min(5, ncol(tab))])

colnames(tab) <- gsub("[\r\n]", "_", colnames(tab))
write.table(tab, file="results/merged.tsv", col.names=T, row.names=F, sep="\t", fileEncoding = "UTF-8")
```

For the final result table, we extracted a total of `r ncol(tab)` different variables from 
the sheets of `r length(unique(tab[,"key_patient"]))` patients. Note again, that for each new visit of the same patient in
the hospital, we have a new data entry, i.e. we have more data entries than we have number 
of patients. Overall, we have `r nrow(tab)` unique data entries.

Below we show the list of variables we extracted from the data:
```{r variables, echo=F}
kable(cbind(colnames(tab)[1:50], 
            c(colnames(tab)[51:ncol(tab)],
              rep("", 11))))
```
## Exploration

We do some basic exploration of the data. For now, we only create four different 
plots, showing the overall numbers of female/male individuals and their
distributions of weight, height and BMI.

```{r exploration, echo=F}
plot_tab <- as.data.frame(tab)
plot_tab[,"weight"] <- as.numeric(as.character(plot_tab[,"weight"]))
plot_tab[,"height"] <- as.numeric(as.character(plot_tab[,"height"]))
plot_tab[,"BMI"] <- as.numeric(as.character(plot_tab[,"BMI"]))


ggplot(aes(x=sex, fill=sex), data=plot_tab) + geom_histogram(stat="count") + 
  ggtitle("Number of female/male patients across all records.")

ggplot(aes(x=weight, fill=sex), data=plot_tab) + geom_histogram() + facet_wrap( ~ sex)+ 
  ggtitle("Distribution of weight across all patiens", "Stratified by sex.")
ggplot(aes(x=height, fill=sex), data=plot_tab) + geom_histogram() + facet_wrap( ~ sex)+ 
  ggtitle("Distribution of height across all patiens", "Stratified by sex.")
ggplot(aes(x=BMI, fill=sex), data=plot_tab) + geom_histogram() + facet_wrap( ~ sex)+ 
  ggtitle("Distribution of BMI across all patiens", "Stratified by sex.")

```

## Conclusion
TODO

## References
TODO